---
title: "WIP Assessing bias corrected data"
author: "Ruth C E Bowyer"
output: 
  github_document:
  date: "`r format(Sys.Date())`"
---


```{r libs and setup, message=FALSE, warning=F}
rm(list=ls())

knitr::opts_knit$set(root.dir="/Volumes/vmfileshare/ClimateData/")

library(terra)
library(sp)
library(exactextractr)
library(dplyr)
library(ggplot2)

dd <- "/Volumes/vmfileshare/ClimateData/"

```


## **0. About**

Assessing the bias correction applied using the ```cmethods``` package. 

## **0. Raw data**

Comparing to the 'Raw' (but reprojected CPM data) -- cropped to Scotland extent (cropping-CPM-to-Scotland.R)

```{r load raw data}

# NEEDS TO BE UPDATED WHEN CAN COPY THE FILES TO THE VM
#  p <- paste0(dd, "Interim/Cropped_UKCPM/")
p <- "~/Library/CloudStorage/OneDrive-TheAlanTuringInstitute/tempdata/"
files <- list.files(p)
raw.files.p <- paste0(p, files)
raw.dat <- lapply(raw.files.p, rast) 

historic.raw <- raw.dat[c(1:7200)]
historic.raw <- rast(historic.raw)
proj.raw <- raw.dat[c(7201:length(raw.dat))]
proj.raw <- rast(proj.raw)
```



##**1. Bias corrected data **

Load and check files 

```{r load data 1}

# Each method is a sep file per decade 
# Create a character vector listing the methods 

p <- paste0(dd, "Debiased/tasmax/")
f <- list.files(p)
i <- c("debiased_delta_method_result", "debiased_linear_scaling_result", "debiased_quantile_delta_mapping", "debiased_quantile_mapping", "debiased_variance_scaling")

# Returns a list of stacked rasters (one for each method) 
debiased_data <- lapply(i, function(x){
  files <- f[grepl(x, f)]
  files.p <- paste0(p, files)
  #Reads in as list of Rasters 
  dat <- lapply(files.p, rast) 
  #Load all together so one raster per method
  files.r <- rast(dat)
})

names(debiased_data) <- i

```

## **2. Assessing the bias correction methods**

### **2a. Time series**

Check the time series over the whole dataset makes sense

#### **2ai. Raw CPM **

Projected time series 

```{r}

df <- as.data.frame(proj.raw, xy=T)


```


#### **2aii. Delta method**

**cemthods** implements Delta Method based on: Beyer, R. and Krapp, M. and Manica, A.: An empirical evaluation of bias correction methods for palaeoclimate simulations (https://doi.org/10.5194/cp-16-1493-2020), which describes the methods as follows:

*The delta method consists of adding the difference between past and present-day simulated climate to present-day observed climate. As such, the delta method assumes that local (i.e. grid-cell-specific) model biases are constant over time (Maraun and Widmann, 2018). For temperature variables (including terrestrial and marine mean annual temperatures and terrestrial temperature of the warmest and coldest months considered here), the bias in a geographical location x is given by the difference between present-day observed and simulated temperature* 

*Precipitation is bounded below by zero and covers different orders of magnitude across different regions. A multiplicative rather than additive bias correction is therefore more common when applying the delta method for precipitation, which corresponds to applying the simulated relative change to the observations (Maraun and Widmann, 2018).*


```{r}
delta <- debiased_data$debiased_delta_method_result
nlyr(delta) #14400 -  correct number for 40Y * (12m*30d) - ie 14400 days

#Create a name col representing the y/m/d - the way the file is reading is not readily useable
ds <- rep(01:30, 12)
ms <- rep(01:12, each=30)
ys <- rep(c(2020:2040, 2060:2080), each=360)
dmy <- paste0(ys,"-",ms,"-",ds)
dmy2<- dmy[which(dmy=="2020-12-1"):which(dmy=="2080-11-30")]
rem <- which(dmy2=="2040-12-1"):which(dmy2=="2060-11-30")
dmy3 <- dmy2[-rem]

names <- paste0("tasmax_", dmy3)
df <- as.data.frame(delta, xy=T)

n2 <- c("x", "y", names)
names(df) <- n2

#Mean and sd of by day calculated for plotting
df.g <- apply(df[c(3:ncol(df))], 2, function(i){
  mean <- mean(i, na.rm=T)
  sd <- sd(i, na.rm=T)
  data.frame(mean=mean, 
             sd.high=mean+sd,
             sd.low=mean-sd)
})

df.g <- df.g %>% purrr::reduce(rbind)
row.names(df.g) <- names
df.g$dmy <- dmy3
```

##### Delta trend daily 2020 - 2040

```{r delta trend 1}
df.g2020 <- df.g[which(df.g$dmy<2060),]

ggplot(df.g2020) + 
  geom_ribbon(aes(x = 1:length(dmy), ymin = sd.low, ymax=sd.high), color="lightgrey", alpha=0.5) +
  geom_line(aes(x=1:length(dmy), y=mean), color="cornflowerblue", group=1) +
  theme_bw() + ylab("Av daily max temp oC") + 
  ggtitle("Delta - tasmax daily") +
 scale_x_discrete(labels = NULL, breaks = NULL) + xlab("Day")

```



##### Delta trend daily 2060 - 2080

```{r delta trend 2}
df.g2060 <- df.g[which(df.g$dmy>2059),]

ggplot(df.g2060) + 
  geom_ribbon(aes(x = 1:length(dmy), ymin = sd.low, ymax=sd.high), color="lightgrey", alpha=0.5) +
  geom_line(aes(x=1:length(dmy), y=mean), color="cornflowerblue", group=1) +
  theme_bw() + ylab("Av daily max temp oC") + 
  ggtitle("Delta - tasmax daily") +
 scale_x_discrete(labels = NULL, breaks = NULL) + xlab("Day")

```

##### Delta trend seasonal averages (full time series)

```{r delta trend seasonal}

#Annotate season based on month index
df.g$season <- ifelse(grepl("-12-|-1-|-2-", df.g$dmy), "Winter",
                      ifelse(grepl("-3-|-4-|-5-", df.g$dmy), "Spring",
                          ifelse(grepl("-6-|-7-|-8-", df.g$dmy), "Summer", "Autumn")))

df.g$year <- as.numeric(sub("-.*", "", df.g$dmy))

#Create a season_year var than considers the same Winter season across 2 years 
## i.e. - Jan 2021 is considered as Winter 2020
df.g$season_year <- ifelse(df.g$season != "Winter"| grepl("-12-", df.g$dmy), 
                           paste0(df.g$season, "_", df.g$year), paste0(df.g$season,"_", df.g$year-1))


#Calculate seasonal mean and SD
seasonal.mean <- df.g %>% 
  group_by(season_year) %>% mutate(mean.seasonal = mean(mean),
                                    sd.high.seasonal = mean.seasonal + sd(mean),
                                    sd.low.seasonal = mean.seasonal - sd(mean))

#Remove daily vals to avoid confusion     
seasonal.mean[c("mean", "sd.high", "sd.low")] <- NULL

#Remove duplicate values
seasonal.mean <- distinct(seasonal.mean, season_year, .keep_all=T) #160 seasons 

```


```{r delta trend seasonal}

#Add in missing years for clearer plotting of trend
dfg_sm <- seasonal.mean
seas.miss <- rep(c("Spring", "Summer", "Autumn", "Winter"), 20)
year.miss <- rep(2041:2060, each=4)
add.s.y <- paste0(seas.miss, "_", year.miss)
add.s.y <- c("Winter_2040", add.s.y)

dfg_sm <- plyr::rbind.fill(dfg_sm,
                           data.frame(year=c(2040, year.miss),
                                      season_year=add.s.y, 
                                      mean.seasonal=NA,
                                      sd.low.seasonal=NA,
                                      sd.high.seasonal=NA))

dfg_sm <- dfg_sm[order(dfg_sm$year),]
```


**Delta - seasonal**

```{r delta seasonal}

ggplot(dfg_sm) + 
  geom_ribbon(aes(x = 1:length(season_year), ymin = sd.low.seasonal, ymax=sd.high.seasonal), color="lightgrey", alpha=0.5) +
  geom_line(aes(x=1:length(season_year), y=mean.seasonal), color="cornflowerblue", group=1) +
  theme_bw() + ylab("Av daily max temp oC") + 
  ggtitle("Delta - tasmax seasonal") + 
  xlab("Season - Year") +
  scale_x_discrete(labels = c(dfg_sm$season_year)) + 
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust=1))

```


**Delta - Winter only**

```{r delta seasonal winter}

dfg_sm_w <- subset(dfg_sm, grepl("Winter", season_year))

ggplot(dfg_sm_w) + 
  geom_ribbon(aes(year, ymin = sd.low.seasonal, ymax=sd.high.seasonal), 
              fill="lightblue3", alpha=0.5) +
  geom_line(aes(year, y=mean.seasonal), color="lightblue4", group=1) +
  theme_bw() + ylab("Av daily max temp oC") + 
  ggtitle("Delta - tasmax seasonal - Winter only") + 
  xlab("Year") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


**Delta - Summer only**

```{r delta seasonal winter}

dfg_sm_s <- subset(dfg_sm, grepl("Summer", season_year))

ggplot(dfg_sm_w) + 
  geom_ribbon(aes(year, ymin = sd.low.seasonal, ymax=sd.high.seasonal), 
              fill="lightblue3", alpha=0.5) +
  geom_line(aes(year, y=mean.seasonal), color="lightblue4", group=1) +
  theme_bw() + ylab("Av daily max temp oC") + 
  ggtitle("Delta - tasmax seasonal - Winter only") + 
  xlab("Year") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```