---
title: "Comparing-Reprojections-HADs"
output: 
  github_document:
    keep_html: true
  html_document:
    theme: cosmo
    toc: TRUE
    toc_float: TRUE
    toc_depth: 4
    df_print: paged
  date: "`r format(Sys.Date())`"
---

```{r libs and setup, message=FALSE, warning=F}
rm(list=ls())

knitr::opts_knit$set(root.dir="/Volumes/vmfileshare/ClimateData/")

library(terra)
library(sp)
library(exactextractr)

dd <- "/Volumes/vmfileshare/ClimateData/"

```


## **0. About**

Bias correction techniques in general require observational data to compare with climate projections in order to appropriately correct the bias. 

The [HadUK grid](https://catalogue.ceda.ac.uk/uuid/bbca3267dc7d4219af484976734c9527) is a 1km x 1km gridded dataset derived from meterological station observations. 

The first UKCP product for review is the UCKP convection-permitting dataset, on a 2.2km grid. Therefore, we are resmapling the 1km grid using bilenear interpolation to 2.2km grid extent. 

We have ran this seperately in both ```r``` and ```python```. The aim of this doc is to:

 - Ensure both methods produce the same result
 - Ensure the grid has been resampled to the correct extent and CRS 

## **1. Data** 

### **1a. HadUK grid resampled in R**
    Resampling script [here](https://github.com/alan-turing-institute/clim-recal/blob/main/R/Resampling.HADs.inR.R)
    The 2.2km grid was derived from a reprojected (to BNG) UKCP 2.2km .nc file


In resampling water vals are resampled as 1.000000e+20  so replacing those vals as NA 

```{r load data 1}

r1 <- paste0(dd,"TestData.R-python/Resampled_HADs_tasmax.2000.01.tif")
r1 <- rast(r1)#Contains 31 layers for each day of Jan

#In the resampling, the method used seemed to have relable all Sea values as '1.000000e+20' so relabelling them here (but to be checked as to why they've been valued like this in the resampling)
r1[r1 > 200] = NA

#check the crs
crs(r1, proj=T)

#Plot to check
plot(r1$tasmax_1)

```

### **1b. HadUK grid resampled in python**
  Resampling script [here](https://github.com/alan-turing-institute/clim-recal/blob/main/python/resampling/resampling_hads.py)
  
**THIS UPDATED 17/02/23**

```{r load data 2}

py.pros.tasmax <- list.files(paste0(dd,"Processed/HadsUKgrid/resampled_2.2km_newgrid/tasmax/day"))
r2 <- py.pros.tasmax[grepl("200001", py.pros.tasmax)] #Same file as resampled above
r2 <- paste0(paste0(dd, "Processed/HadsUKgrid/resampled_2.2km_newgrid/tasmax/day"),"/",r2)
r2 <- rast(r2)
crs(r2, proj=T) #check crs 
## Ok so interesting is missing a crs slot on read - I wonder why this is? This could cause future problem potentially? 

plot(r2$tasmax_1)

```

### **1c. Original HADUK grid**

```{r}
f <- paste0(dd, "Raw/HadsUKgrid/tasmax/day/")
hads.tasmax <- list.files(f)

hads.tasmax2 <- hads.tasmax[grepl("200001", hads.tasmax )] #Same file as resampled above
og <- paste0(f, hads.tasmax2)

og <- rast(og)
crs(og, proj=T)

plot(og$tasmax_1)
```
### **1d. UKCP example**

For comparing the grids

```{r}

f <- paste0(dd,"Processed/UKCP2.2_Reproj/tasmax_bng2/01/latest/tasmax_rcp85_land-cpm_uk_2.2km_01_day_19991201-20001130.tif")
ukcp <- rast(f)
ukcp.r <- ukcp$`tasmax_rcp85_land-cpm_uk_2.2km_01_day_19991201-20001130_31`

crs(ukcp.r, proj=T)

#plot(ukcp.r)
```


### **1e. Cropped extent**
  Just comparing by cropping to Scotland (bbox created [here](https://github.com/alan-turing-institute/clim-recal/tree/main/data/Scotland))

```{r}
scotland <- vect("~/Library/CloudStorage/OneDrive-TheAlanTuringInstitute/CLIM-RECAL/clim-recal/data/Scotland/Scotland.bbox.shp")
```

## **2. Comparisons **

Crop extents to be the same 

```{r}
#Noticed the crop takes longer on r2_c - for investigation!

b <- Sys.time()
r1_c <- terra::crop(r1, scotland, snap="in") 
e <- Sys.time()
e-b

plot(r1_c$tasmax_1)
```

```{r}

b <- Sys.time()
r2_c <- terra::crop(r2, scotland, snap="in") 
e <- Sys.time()
e-b

plot(r2_c$tasmax_1)
```

```{r}
og_c <- terra::crop(og, scotland, snap="in") 
plot(og_c$tasmax_1)
```
Ok there are some differences that I can see from the plot between the two resampled files! 

```{r}
## Cropping to a small area to compare with the same orginal HADS file 
i <- rast()
ext(i) <- c(200000, 210000, 700000, 710000)
```



```{r}
r1_ci <- crop(r1_c, i)
plot(r1_ci$tasmax_1)

#Get number of cells in cropped extent
cells <- cells(r1_ci)

#get coords for all cells (for comparing above)
r.reproj_c_xy <- sapply(cells, function(i){xyFromCell(r1_ci, i)})

r.reproj_c_xy

ext(r1_ci)
```

```{r}

r2_ci <- crop(r2_c, i)
plot(r2_ci$tasmax_1)

ext(r2_ci)
```

```{r}
og_ci <- crop(og_c, i)
ext(og_c)
plot(og_ci$tasmax_1)

```

```{r}
ukcp_c <- terra::crop(ukcp.r, i) 
plot(ukcp_c$`tasmax_rcp85_land-cpm_uk_2.2km_01_day_19991201-20001130_31`)

ext(ukcp_c)
```

```{r}
#Get number of cells in cropped extent
cells <- cells(ukcp_c)

#get coords for all cells (for comparing above)
ukcp_c_xy <- sapply(cells, function(i){xyFromCell(ukcp_c, i)})

ukcp_c_xy
```

```{r}
all(ukcp_c_xy, r.reproj_c_xy)

```

