---
title: "Bias correction assessment of LCAT data"
author: "Ruth Bowyer"
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: cosmo
    toc: TRUE
    toc_float: TRUE
    toc_depth: 4
    code_folding: hide
    df_print: paged
---


```{r libs and setup, message=FALSE, warning=F}
rm(list=ls())

knitr::opts_knit$set(root.dir="/mnt/vmfileshare/ClimateData/")

library(ggplot2)
library(terra)
library(tmap) #pretty maps
library(RColorBrewer)
library(tidyverse)
library(kableExtra)
library(plotrix) #For taylor diagrams 

dd <- "/mnt/vmfileshare/ClimateData/"

```


## **0. About**

LCAT require 'bias corrected' data for the whole of the UK. 
We have applied a widely used approach, quantile mapping, to the data. Specifically, we have used *non-parametric quantile mapping using empirical quantiles* as available in the `Qmap` package. 
Because the data is so large, we have applied this bias correction to the UK broked down into regions, with Scotland brokedn down into further regions (see `R/LCAT/Region.Refs.csv`)

We will now:

- Assess the bias correction using some of the segments
- Process the data back to geotiff 
- Either process as monthly data or as UK wide rasters (maybe just write them seperately) and av across runs

The data is within `ClimateData/Debiased/R/QuantileMapping` and is in RDS format, with each object containing a list.

The objects within this R list are as follows:
- 't.obs': transposed observation df
- 't.cal': transposed calibration df
- 't.proj': transposed projection df (included the validation period)
- 'qm1.hist.a' - bias corrected values for the historical period, values fitted with linear interpolation
- 'qm1.hist.b' - bias corrected values for the historical period, values fitted with tricubic interpolation
- 'qm1.proj.a' - bias corrected values for the validation/projection period, values fitted with linear interpolation
- 'qm1.proj.b' - bias corrected values for the validation/projection period, values fitted with tricubic interpolation

## **1. Bias Correction Assessment: trends**

### **London - tasmax = Run 08**

Using the London region (UKI) as this is the smallest -- not this is the same regional area as the 'three.cities' crops but cut to shapefile edges rather than the square grid 

```{r}

runs <- c("Run05", "Run06", "Run07", "Run08")

London.allruns <- lapply(runs, function(i){
  rds <- paste0(dd,"/Debiased/R/QuantileMapping/resultsL",i,"_UKI_tasmax.RDS")
  readRDS(rds)})

names(London.allruns) <- runs
```

Load in Hads validation data 
(So this can be run for all of the LCAT data, I'm going to read in the whole HADs files for the calibration years)

**The calibration period is 2009-12-01 to 2019-11-30 to relate to the CPM month grouping**

Hads data were also cropped to the regional files for the calibration years - some of the dates might need to be added from the observation (or just be ignored for ease)

```{r}
fp <- paste0(dd, "Interim/HadsUK/Data_as_df/")
f <- list.files(fp)
v <- "tasmax"
reg <- "UKI"
f <- f[grepl("2010_2020", f)&grepl(v,f)&grepl(reg, f)]

obs.val.df <- read.csv(paste0(fp,f)) # Starts here from 2010 - 01 -01 -- because for the BC I removed these vals to align with the cpm years we're missing the month of Dec - so need to update the cpm data to reflect this in the assessment -- wont be a problem for other BC data 

```


### **1b. Check trends**

The next set of chunks visualise the data 

This next chunk converts the dfs back to raster with the correct CRS
```{r convert to df and raster}

## Load a source raster to extract the crs
r <- list.files(paste0(dd, "Reprojected/UKCP2.2/tasmax/05/latest/"))
r <- r[1]
rp <- paste0(dd, "Reprojected/UKCP2.2/tasmax/05/latest/", r)
rast <- rast(rp)

crs <- crs(rast)

## Convert from matrix to df, transpose, create x and y cols - when run in chunk this works fine but for some reason can throw an error when run otherwise
London.df.rL <- lapply(runs, function(i){
  L <- London.allruns[[i]]
    lapply(L, function(x){
    df <- t(x)
    df <- as.data.frame(df)                      
    rn <- row.names(df) #The rownames were saves as x_y coordinates
    xi <- gsub("_.*", "", rn)
    yi <- gsub(".*_", "", rn)
    xy <- data.frame(x = xi, y = yi)
    df <- cbind(xy, df)})
  })

names(London.df.rL) <- runs

## Convert to rasters
London.rasts <- lapply(runs, function(i){
  L <- London.df.rL[[i]]
  lapply(L, function(x){
    r <- rast(x, type="xyz")
    crs(r) <- crs
    return(r)}) 
})

names(London.rasts) <- runs


```

#### **Raster vis comparison**

Random selection of 3 days of the observation, calibration and two adjusted cals, for three historic days
(Note: I'm just plotting the bias corrected with linear interpolation so as to overwhelm with plots)

##### Fig. *Day 1 - 1980-12-01*

```{r, fig.show="hold", out.width="33%"}
tm_shape(London.rasts$Run05$t.obs[[1]]) + tm_raster(title="Observation, 1980-12-01") #Obviously just one call of the observation 
tm_shape(London.rasts$Run05$t.cal[[1]]) + tm_raster(title="Calibration, Run 05, Raw 1980-12-01")
tm_shape(London.rasts$Run06$t.cal[[1]]) + tm_raster(title="Calibration, Run 06, Raw 1980-12-01")
tm_shape(London.rasts$Run07$t.cal[[1]]) + tm_raster(title="Calibration, Run 07, Raw 1980-12-01")
tm_shape(London.rasts$Run08$t.cal[[1]]) + tm_raster(title="Calibration, Run 08, Raw 1980-12-01")
tm_shape(London.rasts$Run05$qm1.hist.a[[1]]) + tm_raster(title="Calibration, Run 05, BC 1980-12-01")
tm_shape(London.rasts$Run06$qm1.hist.a[[1]]) + tm_raster(title="Calibration, Run 06, BC 1980-12-01")
tm_shape(London.rasts$Run07$qm1.hist.a[[1]]) + tm_raster(title="Calibration, Run 07, BC 1980-12-01")
tm_shape(London.rasts$Run08$qm1.hist.a[[1]]) + tm_raster(title="Calibration, Run 08, BC 1980-12-01")


```

##### Fig. *Day 2 - 1991-06-01*

Just to note I was so suprised by how much lower the observation data was for this raster I loaded the raw HADs to check (in resampled_2.2km/tasmax and the original 1km grid it does reflect it - it just seems very low)

```{r, fig.show="hold", out.width="33%"}
tm_shape(London.rasts$Run05$t.obs[[3781]]) + tm_raster(title="Observation, 1991-06-01") #Obviously just one call of the observation 
tm_shape(London.rasts$Run05$t.cal[[3781]]) + tm_raster(title="Calibration, Run 05, Raw 1991-06-01")
tm_shape(London.rasts$Run06$t.cal[[3781]]) + tm_raster(title="Calibration, Run 06, Raw 1991-06-01")
tm_shape(London.rasts$Run07$t.cal[[3781]]) + tm_raster(title="Calibration, Run 07, Raw 1991-06-01")
tm_shape(London.rasts$Run08$t.cal[[3781]]) + tm_raster(title="Calibration, Run 08, Raw 1991-06-01")
tm_shape(London.rasts$Run05$qm1.hist.a[[3781]]) + tm_raster(title="Calibration, Run 05, BC 1991-06-01")
tm_shape(London.rasts$Run06$qm1.hist.a[[3781]]) + tm_raster(title="Calibration, Run 06, BC 1991-06-01")
tm_shape(London.rasts$Run07$qm1.hist.a[[3781]]) + tm_raster(title="Calibration, Run 07, BC 1991-06-01")
tm_shape(London.rasts$Run08$qm1.hist.a[[3781]]) + tm_raster(title="Calibration, Run 08, BC 1991-06-01")

```



##### Fig. *Day 3 - 2000-08-01*


```{r, fig.show="hold", out.width="33%"}
tm_shape(London.rasts$Run05$t.obs[[7081]]) + tm_raster(title="Observation, 2000-08-01") #Obviously just one call of the observation 
tm_shape(London.rasts$Run05$t.cal[[7081]]) + tm_raster(title="Calibration, Run 05, Raw 2000-08-01")
tm_shape(London.rasts$Run06$t.cal[[7081]]) + tm_raster(title="Calibration, Run 06, Raw 2000-08-01")
tm_shape(London.rasts$Run07$t.cal[[7081]]) + tm_raster(title="Calibration, Run 07, Raw 2000-08-01")
tm_shape(London.rasts$Run08$t.cal[[7081]]) + tm_raster(title="Calibration, Run 08, Raw 2000-08-01")
tm_shape(London.rasts$Run05$qm1.hist.a[[7081]]) + tm_raster(title="Calibration, Run 05, BC 2000-08-01")
tm_shape(London.rasts$Run06$qm1.hist.a[[7081]]) + tm_raster(title="Calibration, Run 06, BC 2000-08-01")
tm_shape(London.rasts$Run07$qm1.hist.a[[7081]]) + tm_raster(title="Calibration, Run 07, BC 2000-08-01")
tm_shape(London.rasts$Run08$qm1.hist.a[[7081]]) + tm_raster(title="Calibration, Run 08, BC 2000-08-01")

```

#### **Calibration period - annual trends**


```{r}
#Returns a list of dfs in handy format for graphing
London.dfg.rL <- lapply(runs, function(i){
  L <- London.df.rL[[i]]
  names(L)[1:3] <- c("obs", "cal", "proj") 
    dfg <- lapply(names(L), function(ii){
      dfi <- L[[ii]]
      x <- 3:ncol(dfi) #ignore cols 1 & 2 with x y
      #Calc mean and sd
      dfx <- lapply(x, function(x){
        y <- dfi[,x]
        mean <- mean(y, na.rm=T)
        sd <- sd(y, na.rm=T)
        dfr <- data.frame(mean=mean, 
             sd.high=mean+sd,
             sd.low=mean-sd)
        names(dfr) <- paste0(ii, ".", names(dfr))
        dfr$day <- names(dfi)[x]
        return(dfr)
      })

      dfx_g <- dfx %>% purrr::reduce(rbind)
    })

    names(dfg) <- c("obs.daymeans", "raw.cal.daymeans",
                       "raw.proj.daymeans", "bc.a.cal.daymeans",
                       "bc.b.cal.daymeans", "bc.a.proj.daymeans",
                       "bc.b.proj.daymeans")
    
    return(dfg)
})

names(London.dfg.rL) <- runs
```


```{r}
#Create a df for all of the runs to plot
##Add a day index to align the cal and obs 

London.dfg.calp.L <- lapply(runs, function(i){
   dfg <- London.dfg.rL[[i]]
  dfg.calp <- dfg[c("obs.daymeans", "raw.cal.daymeans",
                       "bc.b.cal.daymeans", "bc.a.cal.daymeans")]

    dfg.calp <- lapply(dfg.calp, function(x){
    x$dayi <- 1:nrow(x)
    x$day<- NULL
    return(x)
    })
    
    
  dfg.calp <- dfg.calp %>% reduce(merge, "dayi")
  dfg.calp$Run <- i
  return(dfg.calp)})

names(London.dfg.calp.L) <- runs

London.dfg.calp <- London.dfg.calp.L %>% reduce(rbind)

```

```{r}

London.dfg.calp_m <- reshape2::melt(London.dfg.calp, id=c("dayi", "Run")) #create long df for plotting multiple lines

London.dfg.calp_mm <- London.dfg.calp_m[grepl("mean", London.dfg.calp_m$variable),] #For easy vis, only keep mean vals
```

#### Fig. Calibration period - annual mean

```{r Historic trend 1}

ggplot(London.dfg.calp_mm, aes(dayi, value, group=variable, colour=variable)) + 
  geom_line(alpha=0.7) +
  facet_wrap(.~Run) +
  theme_bw() + ylab("Av daily max temp oC") + 
  ggtitle("Tasmax Hisotric trends") +
 scale_x_discrete(labels = NULL, breaks = NULL) + xlab("Day, 1980.12.01 - 2009.12.01") +
  scale_color_brewer(palette="Set1", name="Model", labels=c("Obs (Hads)", "Raw CPM", "BC CPM 1", "BC CPM 2"))

```

#### **Seasonal trends - Calibration period **

Annotate season based on month index - the dates have different formats depending on the input data (ie hads vs cpm) so am pulling out the necessary to adjust sep 

```{r}

seasonal.means <- lapply(runs, function(r){
  dfg <- London.dfg.rL[[r]]
    #Hads/obs df
    obs.daymeans.df <- dfg$obs.daymeans

      x <- obs.daymeans.df$day
      obs.daymeans.df$season <- ifelse(grepl("1231_|0131_|0228_|0229_", x),
                                       "Winter",
                      ifelse(grepl("0331_|0430_|0531_", x), "Spring",
                          ifelse(grepl("0630_|0731_|0831_", x), "Summer", "Autumn")))

#Note: the seasons should each have 90 days but seemingly Winter and Autumn have 89 and Spring and Summer have 91 - this is due to how the manual aligning worked out and should be updated when the hads data is re-run 

    #Create season_year - All Winter months apart from Dec to be added to the previous year (ie     Winter 2000) would be the Dec of 2000 to the Feb of 2001
    year <- gsub("^[^_]*_", "", x)
    year <- as.numeric(substr(year, 1,4))
    obs.daymeans.df$season_year <- ifelse(grepl("0131_|0228_|0229_", x), 
                                      paste0(year-1, obs.daymeans.df$season), 
                                      paste0(year, obs.daymeans.df$season))
    # Mutate to a seasonal mean df 
    obs.seasonal.mean.df <- aggregate(obs.daymeans.df[[1]], list(obs.daymeans.df[["season_year"]]), function(x) c(seasonal.mean = mean(x), sd.high.seasonal = mean(x) + sd(x), sd.low.seasonal = mean(x) - sd(x)))
    obs.seasonal.mean.df<- data.frame(season_year=obs.seasonal.mean.df$Group.1,
                                        seasonal.mean=obs.seasonal.mean.df$x[,"seasonal.mean"],
                                        sd.high.seasonal = obs.seasonal.mean.df$x[,"sd.high.seasonal"],
                                        sd.low.seasonal = obs.seasonal.mean.df$x[,"sd.low.seasonal"])
    

  #Grouping variable for later vars 
  obs.seasonal.mean.df$model <- "obs"
   
   
    dfg.seasonal.mean <- lapply(c("raw.cal.daymeans", "bc.b.cal.daymeans",
                                         "bc.a.cal.daymeans"), function(i){
        df <- dfg[[i]]
        x <- df$day
        x <- gsub(".*_", "", x)
        x <- as.numeric(x)
      #The CPM days are consecutive 1 - 360 by year
        df$season <- ifelse(x<91, "Winter",
                      ifelse(x<181, "Spring",
                          ifelse(x<271, "Summer", "Autumn")))
  
  #Create season_year - All Winter months apart from Dec to be added to the previous year (ie     Winter 2000) would be the Dec of 2000 to the Feb of 2001  
        year <- gsub(".*day_", "", df$day)
        year <- as.numeric(substr(year, 1,4))
        df$season_year <- ifelse(x>29&x<91, 
                         paste0(year-1, df$season), 
                         paste0(year, df$season))
  
    # Mutate to a seasonal mean -- cant get this to run in tidyverse within loop as cant seem to get col indexing working so:
       df2 <- aggregate(df[[1]], list(df[["season_year"]]), function(x) c(seasonal.mean = mean(x), sd.high.seasonal = mean(x) + sd(x), sd.low.seasonal = mean(x) - sd(x)))
       
       df2 <-    data.frame(season_year=df2$Group.1,
                                        seasonal.mean=df2$x[,"seasonal.mean"],
                                        sd.high.seasonal = df2$x[,"sd.high.seasonal"],
                                        sd.low.seasonal = df2$x[,"sd.low.seasonal"])

        df2$model <- gsub(".daymeans","",i)

       return(df2)})

  dff <- c(list(obs.seasonal.mean.df), dfg.seasonal.mean) %>% reduce(rbind)  
  dff$Run <- r
  return(dff)
})

names(seasonal.means) <- runs

seasonal.means.df <- seasonal.means %>% reduce(rbind)

```

#### Fig. Calibration period - seasonal mean

```{r}

ggplot(seasonal.means.df, aes(season_year, seasonal.mean, group=model, colour=model)) + 
  geom_line() +
  facet_wrap(.~Run) +
  theme_bw() + ylab("Av daily max temp oC") + 
  ggtitle("Tasmax Hisotric trends") +
 scale_x_discrete(labels = NULL, breaks = NULL) + xlab("Seasonal averages, 1980.12.01 - 2009.12.01") +
  scale_color_brewer(palette="Set1", name="Model")

```


##### *Summer only*

```{r Raw seasonal winter}

dfg_sm<- subset(seasonal.means.df, grepl("Summer", season_year))

ggplot(dfg_sm, aes(season_year, seasonal.mean, group=model, colour=model)) + 
  geom_line(alpha=0.7) +
  facet_wrap(.~Run) +
  theme_bw() + ylab("Av daily max temp oC -Summer average") + 
  ggtitle("Tasmax Hisotric trends") +
 scale_x_discrete(labels = NULL, breaks = NULL) + xlab("Summer averages, 1980.12.01 - 2009.12.01") +
  scale_color_brewer(palette="Set1", name="Model")

```
It looks purple because the two bc methods aren't revealing much difference so subsetting to just one instead


```{r}

dfg_sm<- subset(seasonal.means.df, !grepl("bc.b.cal", model)&grepl("Summer", season_year))

ggplot(dfg_sm, aes(season_year, seasonal.mean, group=model, colour=model)) + 
  geom_line(alpha=0.7) +
  facet_wrap(.~Run) +
  theme_bw() + ylab("Av daily max temp oC -Summer average") + 
  ggtitle("Tasmax Hisotric trends") +
 scale_x_discrete(labels = NULL, breaks = NULL) + xlab("Seasonal averages, 1980.12.01 - 2009.12.01") +
  scale_color_brewer(palette="Set1", name="Model")

```

#### *Annual trends - seasonal max*

For tasmax - grouping to season and calculating the seasonal maxima vals (i.e. rather than means above) 

```{r}

#Convert to max, out put a df in easy fig format 
London.dfg.max <- lapply(runs, function(r){
  L <- London.df.rL[[r]]
  names(L)[1:3] <- c("obs", "cal", "proj") 
    dfg <- lapply(names(L), function(ii){
      dfi <- L[[ii]]
      x <- 3:ncol(dfi) #ignore cols 1 & 2 with x y
      #Calc maxima of the 
      dfx <- lapply(x, function(x){
        xx <- dfi[,x]
        data.frame(max=max(xx, na.rm=T), day= names(dfi)[x])
      })

      dfx_g <- dfx %>% purrr::reduce(rbind)
    })

    names(dfg) <- paste0(names(L), ".max")
    return(dfg)
})

names(London.dfg.max) <- runs

seasonal.max.cal <- lapply(runs, function(r){
  dfg <- London.dfg.max[[r]]
    #Hads/obs df
    df1 <- dfg$obs.max
      x <- df1$day
      df1$season <- ifelse(grepl("1231_|0131_|0228_|0229_", x),
                                       "Winter",
                      ifelse(grepl("0331_|0430_|0531_", x), "Spring",
                          ifelse(grepl("0630_|0731_|0831_", x), "Summer", "Autumn")))

#Note: the seasons should each have 90 days but seemingly Winter and Autumn have 89 and Spring and Summer have 91 - this is due to how the manual aligning worked out and should be updated when the hads data is re-run 

    #Create season_year - All Winter months apart from Dec to be added to the previous year (ie     Winter 2000) would be the Dec of 2000 to the Feb of 2001
    year <- gsub("^[^_]*_", "", x)
    year <- as.numeric(substr(year, 1,4))
    df1$season_year <- ifelse(grepl("0131_|0228_|0229_", x), 
                                      paste0(year-1, df1$season), 
                                      paste0(year, df1$season))
    # Mutate to a seasonal mean df 
    obs.seasonal.max.df <- aggregate(df1[[1]], list(df1[["season_year"]]), max)
  #Grouping variable for later vars 
  obs.seasonal.max.df$model <- "obs"
   
    dfg.seasonal.max <- lapply(c("cal.max", "qm1.hist.a.max",
                                         "qm1.hist.b.max"), function(i){
        df <- dfg[[i]]
        x <- df$day
        x <- gsub(".*_", "", x)
        x <- as.numeric(x)
      #The CPM days are consecutive 1 - 360 by year
        df$season <- ifelse(x<91, "Winter",
                      ifelse(x<181, "Spring",
                          ifelse(x<271, "Summer", "Autumn")))
  
  #Create season_year - All Winter months apart from Dec to be added to the previous year (ie     Winter 2000) would be the Dec of 2000 to the Feb of 2001  
        year <- gsub(".*day_", "", df$day)
        year <- as.numeric(substr(year, 1,4))
        df$season_year <- ifelse(x>29&x<91, 
                         paste0(year-1, df$season), 
                         paste0(year, df$season))
  
    # Mutate to a seasonal mean -- cant get this to run in tidyverse within loop as cant seem to get col indexing working so:
       df2 <- aggregate(df[[1]], list(df[["season_year"]]), max)

        df2$model <- gsub(".max","",i)

       return(df2)})

  dff <- c(list(obs.seasonal.max.df), dfg.seasonal.max) %>% reduce(rbind)  
  dff$Run <- r
  return(dff)
})

names(seasonal.max.cal) <- runs

seasonal.maxima.df <- seasonal.max.cal %>% reduce(rbind)
names(seasonal.maxima.df) <- c("season_year", "max", "model", "Run")
```

#### Fig. Calibration period - seasonal max

```{r}

ggplot(seasonal.maxima.df, aes(season_year, max, group=model, colour=model)) + 
  geom_line() +
  facet_wrap(.~Run) +
  theme_bw() + ylab("Max daily max temp oC") + 
  ggtitle("Tasmax Hisotric trends") +
 scale_x_discrete(labels = NULL, breaks = NULL) + xlab("Seasonal averages, 1980.12.01 - 2009.12.01") +
  scale_color_brewer(palette="Set1", name="Model")

```

#### Fig. Calibration period -  *Summer only*

```{r}

dfg_sm<- subset(seasonal.maxima.df, !grepl("qm1.hist.b", model)&grepl("Summer", season_year))

ggplot(dfg_sm, aes(season_year, max, group=model, colour=model)) + 
  geom_line(alpha=0.7) +
  facet_wrap(.~Run) +
  theme_bw() + ylab("Av daily max temp oC -Summer average") + 
  ggtitle("Tasmax Historic trends") +
 scale_x_discrete(labels = NULL, breaks = NULL) + xlab("Seasonal Summer averages, 1980.12.01 - 2009.12.01") +
  scale_color_brewer(palette="Set1", name="Model")

```

#### Create validaton df list

Adding in the observational HADs data and aligning based on dates

*Note* So as not to re-run the UK wide LCAT data processing, a workaround was added to the bias correction function used to group the obs data - this means that to align the validation cpm data we have to remove a month in the beginning 

```{r}

#Extract validation period of raw and bias corrected CPM data
val.dfs <- lapply(runs, function(r){
    London.df <- London.df.rL[[r]]
    cpm.val.dfs <- lapply(London.df[c("t.proj", "qm1.proj.a", "qm1.proj.b")], function(x){
    i <- grep("20191201-20201130_30", names(x))[1]
    df <- x[,1:i]
  })

    #Using the old cpm data for the hads obs - so need to remove the dates to ensure theres 30 days per year
    remove <- c("0229_29", "0430_30", "0731_31", "0930_30", "1130_30")
    remove <- paste0(remove, collapse = "|")
      
    obs.val.df <- obs.val.df[,!grepl(remove, names(obs.val.df))]
    row.names(obs.val.df) <- paste0(obs.val.df$x, "_", obs.val.df$y)

    val.dfs <- c(list(obs.val.df), cpm.val.dfs)
    names(val.dfs) <- c("obs.val.df", "raw.cpm.val", "bc1.cpm.val", "bc2.cpm.val")
    return(val.dfs)
    })

names(val.dfs) <- runs
```


#### *Validation period - annual trends - seasonal mean*

(To be added)

#### *Validation period - annual trends - seasonal max*

(To be added)

## **2. Bias Correction Assessment: Metrics**

Using the validation data set for this


```{r}
#Convert dfs to a vector
val.dfs.v <- lapply(runs, function(r){
  dfs <- val.dfs[[r]]
  dfs2 <- lapply(dfs, function(d){
  #Remove x and y 
  d$x <- NULL
  d$y <- NULL
  #Convert to single vector
  unlist(as.vector(d))})
  names(dfs2) <- names(dfs)

val.dfs.v.df <- dfs2 %>% reduce(cbind)
val.dfs.v.df <- as.data.frame(val.dfs.v.df)})

names(val.dfs.v) <- runs
```

```{r}
val.dfs.v <- lapply(runs, function(r){
  df <- val.dfs.v[[r]]
  names(df) <-paste0(r, ".", c("obs.val.df", "raw.cpm.val", "bc1.cpm.val", "bc2.cpm.val"))
  return(df)
})

#Convert to a single df 
val.dfs.v.allruns <- val.dfs.v %>% reduce(cbind)

#Remove duplicate obs (pulled through across each run)
val.dfs.v.allruns[c("Run06.obs.val.df", "Run07.obs.val.df", "Run08.obs.val.df")] <- NULL
names(val.dfs.v.allruns)[1] <- "obs.val"
```

### **2a. Descriptive statistics**

```{r descriptives validation}

descriptives <- apply(val.dfs.v.allruns,2, function(x){ 
  per <- data.frame(as.list(quantile(x, probs=c(0.1, 0.9))))
  data.frame(mean=mean(x), sd=sd(x), min = min(x), per10th=per$X10.,per90th=per$X90., max = max(x))
})

descriptives <- descriptives %>% reduce(rbind)
row.names(descriptives) <- names(val.dfs.v.allruns)
t(descriptives)
```



#### **Distribution**

```{r}

names(val.dfs.v) <- runs
val.dfs.v_fordist <- lapply(runs, function(r){
  df <- val.dfs.v[[r]]
  names(df) <- c("obs", "raw.cpm", "bc1.cpm", "bc2.cpm")
  df$run <- paste0(r)
  return(df)
})

#Convert to a single df 
val.dfs.v.allruns_fordist <- val.dfs.v_fordist %>% reduce(rbind)
val.dfg <- reshape2::melt(val.dfs.v.allruns_fordist, id="run")
```

#### Fig.Density plot of validation period 

```{r}
ggplot(subset(val.dfg, variable!="bc2.cpm"), aes(value, fill=variable, colour=variable)) + 
  geom_density(alpha = 0.3, position="identity") + 
 facet_wrap(.~ run) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1")

```
### **2b. Model fit statistics**

Using the following to assess overall fit: 

- **R-squared (rsq)**
- **Root Square Mean Error (RMSE)**
- **Nash-Sutcliffe Efficiency (NSE):** Magnitude of residual variance compared to measured data variance, ranges -∞ to 1, 1 = perfect match to observations
- **Percent bias (PBIAS):** The optimal value of PBIAS is 0.0, with low-magnitude values indicating accurate model simulation. Positive values indicate overestimation bias, whereas negative values indicate model underestimation bias.

```{r rsq}
actual <- val.dfs.v.allruns$obs.val

rsq <- sapply(val.dfs.v.allruns[c(2:ncol(val.dfs.v.allruns))], function(x){
  cor(actual, x)^2
})

 t(data.frame(as.list(rsq), row.names = "RSQ"))
```

```{r rmse}

rmse <- sapply(val.dfs.v.allruns[c(2:ncol(val.dfs.v.allruns))], function(x){
  sqrt(mean((actual - x)^2))
})

```

```{r pbias}

pbias <- sapply(val.dfs.v.allruns[c(2:ncol(val.dfs.v.allruns))], function(x){
  hydroGOF::pbias(x, actual)
})

```

```{r nse}
nse <- sapply(val.dfs.v.allruns[c(2:ncol(val.dfs.v.allruns))], function(x){
  hydroGOF::NSE(x, actual)
})

```

Highlighting the bias corrected statistics 

```{r pretty kable}

k <- cbind(rsq, rmse, pbias, nse)
k %>% 
  kable(booktabs = T) %>%
  kable_styling() %>%
  row_spec(grep(".bc.",row.names(k)), background = "lightgrey")

```



## **3. Bias Correction Assessment: Metric specific - tasmax**

### **3b Days above 30 degrees**

(Not considered consecutively here)

```{r}
val.dfs.v.allruns$year <- substr(row.names(val.dfs.v.allruns), 8,11)

over30 <- lapply(names(val.dfs.v.allruns), function(i){
  x <- val.dfs.v.allruns[,i]
  df <- aggregate(x, list(val.dfs.v.allruns$year), function(x){sum(x>=30)})
  names(df) <- c("year", paste0("Days.over.30.", i))
                 return(df)
})

over30 %>% reduce(left_join, "year")
```


### **Number of heatwaves per annum**

(to be added)

#### **For future work**

The number of quantiles selected will effect the efficacy of the bias correction: lots of options therefore with this specific method


