---
title: "Quantile Mapping across three cities"
author: "Ruth C E Bowyer"
date: "`r format(Sys.Date())`"
output:
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 0. About

Testing `Qmap` for the 3 cities 

```{r libraries dd}
rm(list=ls())

library(qmap)
library(terra)
library(tidyverse) 
library(doParallel)

dd <- "/mnt/vmfileshare/ClimateData/"

```

## Apply bias correction by variable/run in `qmap`

`qmap` offers a few different bias correction options. 
The below chunks calls a function that wraps the qmap function to loop over the cities and apply the bias correction
It returns a list object, where for each model, we have:
- a matrix of the calibration observation period (01-12-1980 to 30-11-2010 )
- a matrix of values relating to the validation obs period (hads data 01-12-2010 to 01-12-2020)
the raw values and the adjusted values for the CPM for the 
"t.obs", "val.df", "t.cal", "t.proj", "qm1.hist", "qm1.val.proj"

Data has been pre-processed from cropped using 'converting_city_crops_to_df.R' to .csv 

##1. Empirical Quantile Mapping 



```{r warning = F}
setwd("/home/dyme/Desktop/clim-recal/clim-recal")
source("R/bias-correction-methods/apply_qmapQuant_to_crpd_df_fn.R")

cities <- c("London", "Glasgow", "Manchester")
run <- c("Run05", "Run06", "Run07","Run08")
var <- c("tasmax", "tasmin", "pr")

lapply(cities, function(x){
  rd <- paste0(dd, "Debiased/R/QuantileMapping/three.cities/", x, "/")
  
  apply_qmapQUANT_to_cropped_df(region = x,
                                var = var,
                                Runs = run,
                                pd = paste0(dd, "Interim/CPM/Data_as_df/three.cities/"),#Parent directory where dataframes of cpm data are located
                                pd.obs = paste0(dd, "Interim/HadsUK/Data_as_df/three.cities/"),#Parent directory where dataframes of obs data are
                              val.startdate = "20101201", #The first date of the validation period. eg 20101201 All dates before this time will be taken as the calibration
                              ## These args to be passed to qmapQUANT itself:
                                     qstep = 0.1, # numeric value between 0 and 1, e.g 0.1. The quantile mapping is fitted only for the quantiles defined by quantile(0,1,probs=seq(0,1,by=qstep).
                                     nboot = 1, #numeric value 1 or greater - nboot number of bootstrap samples used for estimation of the observed quantiles. If nboot==1 the estimation is based on all (and not resampled) data.

                                     type = "linear", #interpolation method to use for fitting data to the predictions )(eg linear, tricubic)
                              rd = rd)
})
              
                  
```


```{r}

```