---
title: "WIP MBC in R"
author: "Ruth C E Bowyer"
date: '2023-02-28'
output: 
  github_document:
    keep_html: true
  html_document:
    theme: cosmo
    toc: TRUE
    toc_float: TRUE
    toc_depth: 4
    df_print: paged
  date: "`r format(Sys.Date())`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 0. About

Testing Bias Correction methods from the MBC package in R

- For univariate methods, just replicating for a subset the same methods applied in python
- For multivariate methods, will try on all geographical locations 

```{r libraries dd}
rm(list=ls())

library(MBC)
library(terra)
library(sf)
library(exactextractr)

#Loaded package versions
x <- c("MBC", "terra", "sf", "exactextractr")
lapply(x,packageVersion)

dd <- "/Volumes/vmfileshare/ClimateData/"


```


## Observational data

Reprojected HADs data 

First going to subset to a specific area 

```{r}
### Observational HADs
 HADbrickL_crop_Y_2.2km$Y1988$X1988.02.16.12.00.00

#Subset to first 360 days per year 
 HADbrickL_crop_Y_2.2km360 <- lapply(HADbrickL_crop_Y_2.2km, function(x){
   x[[1:360]]
 })
```

```{r}
n <- names(UKCP_R2)[1:20]

# Returns a list of dataframes for each year by rows of the raster cells
ObsL <- lapply(n, function(i){
    Sp <- HADbrickL_crop_Y_2.2km360[[i]]
    Obs <- as(Sp, 'SpatialPolygonsDataFrame')
    Obs <- Obs@data
  })

#Reduce to one df
Obs_df <- ObsL %>% reduce(cbind)
Obs_oc <-  unlist(Obs_df)
```

##### Model outputs - calibration

```{r}
### Rename the Model years 
#First running on one run to test and then will loop
UKCP_R2 <- Lewisham.Bricks.AllYears$Run02

names(UKCP_R2) <- paste0("Y",c(1981:2000, 2021:2040, 2061:2080))

n <- names(UKCP_R2)[1:20]

# Returns a list of dataframes for each year by rows of the raster cells
CalL <- lapply(n, function(i){
    Sp <- UKCP_R2[[i]]
    Cal <- as(Sp, 'SpatialPolygonsDataFrame')
    Cal <- Cal@data
  })

#Reduce to one df
Cal_df <- CalL %>% reduce(cbind)
Cal_oc <- unlist(Cal_df)
```

##### Model outputs - projection

```{r}
### Rename the Model years 
#First running on one run to test and then will loop

n <- names(UKCP_R2)[21:60]

# Returns a list of dataframes for each year by rows of the raster cells
ProjL <- lapply(n, function(i){
    Sp <- UKCP_R2[[i]]
    Proj <- as(Sp, 'SpatialPolygonsDataFrame')
    Proj <- Proj@data
  })

#Reduce to one df
Proj_df <- ProjL %>% reduce(cbind)
Proj_oc <- unlist(Proj_df)
```


#### Univariate quantile mapping

This package for univar mapping but soon to be multivar

Following the vignette here (first using the test data) https://cran.r-project.org/web/packages/MBC/MBC.pdf 

Do this seperately for each run

Below method is univariate quanitle mapping

Next step multivariate mapping in same package 

```{r}
library(MBC)

# Univariate
  fit.qdm <- 
    QDM(o.c=Obs_oc, #vector of observed samples during the calibration period.
        m.c=Cal_oc, #vector of model outputs during the calibration period.
        m.p=Proj_oc, #vector of model outputs during the projected period.
        ratio=FALSE, #logical value indicating if samples are of a ratio quantity (e.g., precipitation) -- false as is temp
        trace=Inf) #numeric value indicating the threshold below which values of a ratio quantity (e.g., ratio=TRUE) should be considered exact zeros. -- need to read up on this 

  
mhat.c  <- fit.qdm$mhat.c
mhat.p   <- fit.qdm$mhat.p  
  
```

-- To do -- 

Next repopulate the dataframe as above, add back to the dataslots, and compare!
Then write in a loop
Then see how long it takes 
Then think about the next way of doing it