---
title: "**Identifying Runs for bias correction**"
author: "Ruth C E Bowyer"
date: "`r format(Sys.Date())`"
output:
  github_document
---

```{r libs and setup, message=FALSE, warning=F}
rm(list=ls())

library(tidyverse)
library(ggplot2)
library(RColorBrewer)

```


## **0. About**

Script to identify the mean, 2nd highest and 2nd lowers daily tasmax per UKCP18 CPM run.

These runs will be the focus of initial bias correction focus

## **1. Load Data**

Data is tasmax runs converted to dataframe using sript 'ConvertingAllCPMdataTOdf.R', with files later renamed.Then daily means for historical periods and future periods were calculated using 'calc.mean.sd.daily.R' and summaries saved as .csv

In retrospect the conversion to df might not have been necessary/the most resource efficient, see comment here:https://tmieno2.github.io/R-as-GIS-for-Economists/turning-a-raster-object-into-a-data-frame.html -- this was tested and using `terra::global` to calculate the raster-wide mean was less efficient

**Update 13.05.23** - Adding in infill data, mean to be calculated over the whole time period

As of June 2023, the tasmax-as-dataframe and tasmax daily means and the df data is located in `vmfileshare/Interim/tasmax_dfs/`

There is an error in the naming convention - Y00_Y20 should be Y01 to reflect the infill data time period (although this does cover a breif period of 2000) - to be updated in future


```{r}

Runs <- c("01", "04", "05", "06", "07", "08", "09", "10", "11", "12", "13", "15")

files <- list.files("/Users/rbowyer/Library/CloudStorage/OneDrive-TheAlanTuringInstitute/tempdata/")

files <- files[grepl(".csv", files)]
fp <- paste0("/Users/rbowyer/Library/CloudStorage/OneDrive-TheAlanTuringInstitute/tempdata/", files)
```

```{r}

# Creating objects for names and filepath for each of the timer periods, for easy loading
names <- gsub("df.avs_|.csv|df.", "", files)
i <- c("hist", "Y00_Y20","Y21_Y40", "Y41_Y60", "Y61_Y80")

namesL <- lapply(i, function(i){
  n <- names[grepl(i, names)]
  })

names(namesL) <- paste0("names_",i)
list2env(namesL, .GlobalEnv)
```

```{r load data}

dfL <- lapply(i, function(i){
  fp <- fp[grepl(i, fp)]
  dfs <- lapply(fp, read.csv)
  n <- namesL[[paste0("names_",i)]]
  names(dfs) <- n
  return(dfs)
  })

names(dfL) <- paste0("dfs_", i)
list2env(dfL, .GlobalEnv)
```

## **2. Comparing Runs**

### **2a. Historical figures**

```{r}

Y <- rep(c(1981:2000), each=360)

dfs_hist <- lapply(names_hist, function(i){
  df <- dfs_hist[[i]]
  names(df) <- c("day", "mean", "sd")
  df$model <- i
  df$dn <- 1:nrow(df)
  df$Y <- Y
  return(df)
})

#Create a single df in long form of Runs for the historical period
historical_means <- dfs_hist %>% reduce(rbind)

```

### **Time series - daily**

```{r fig.height=8}

ggplot(historical_means) +
    geom_line(aes(x=dn, y=mean, group=model, colour=model)) +

    theme_bw() + xlab("Day (Historical 1980 - 2000)") +
    ylab("Daily mean max temp (tasmax) oC") +

   scale_colour_brewer(palette = "Paired", name = "") +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "none") +
  facet_wrap(.~ model, ncol=3)


```


### **boxplot - mean historical**

```{r}
#Create a pallete specific to the runs so when reordered maintain the same colours
historical_means$model <- as.factor(historical_means$model)
c <- brewer.pal(12, "Paired")
my_colours <- setNames(c, levels(historical_means$model))

```


```{r}

historical_means %>%
  mutate(model = fct_reorder(model, mean, .fun='median')) %>%
    ggplot(aes(x=reorder(model, mean), y=mean, fill=model)) +
    geom_boxplot() + theme_bw() +
    ylab("Mean daily max temp (tasmax) oC") + xlab("model") +
    scale_fill_manual(values = my_colours) +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```

### **qqplot - daily means**

```{r fig.height=8}

ggplot(historical_means, aes(sample=mean, colour=factor(model))) +
  stat_qq() +
  stat_qq_line()+
  theme_bw()+
  scale_color_manual(values = my_colours) +
  facet_wrap(.~model, ncol=3)

```


### **Time series - annual mean**

```{r message=FALSE}

#Aggregating to year for annual average

historical_means$Yf <- as.factor(historical_means$Y)

historical_means_y <- historical_means %>%
  group_by(Yf, model) %>%
  dplyr::summarise(mean.annual=mean(mean, na.rm=T), sd.annual=sd(mean, na.rm = T))
```

```{r}
ggplot(historical_means_y) +
    geom_line(aes(x = as.numeric(Yf), y=mean.annual,
              color=model)) +
    theme_bw() + xlab("Year (Historical 1980 - 2000)") +
    ylab("Annual mean of mean daily max temp (tasmax) oC") +
  scale_fill_brewer(palette = "Paired", name = "") +
   scale_colour_brewer(palette = "Paired", name = "") +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```


```{r}

# Plotting with SDs in geom_ribbon to see if anything wildely different
ggplot(historical_means_y) +
    geom_ribbon(aes(as.numeric(Yf), y=mean.annual,
                               ymin = mean.annual - sd.annual,
                              ymax= mean.annual + sd.annual,
                    fill=model),  alpha=0.4) +
    geom_line(aes(x = as.numeric(Yf), y=mean.annual,
              color=model)) +
    theme_bw() + xlab("Year (Historical 1980 - 2000)") +
    ylab("Annual mean of mean daily max temp (tasmax) oC") +
  scale_fill_brewer(palette = "Paired", name = "") +
   scale_colour_brewer(palette = "Paired", name = "") +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + facet_wrap(.~model, ncol=3)

```

### **boxplot - annual mean historical**

```{r}

historical_means_y %>%
  mutate(model = fct_reorder(model, mean.annual, .fun='median')) %>%
    ggplot(aes(x=reorder(model, mean.annual), y=mean.annual, fill=model)) +
    geom_boxplot() + theme_bw() +
    ylab("Annual max daily max temp oC") + xlab("model") +
   scale_fill_manual(values = my_colours) +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```


### **qqplot - annual means**

```{r fig.height=8}

ggplot(historical_means_y, aes(sample=mean.annual, colour=factor(model))) +
  stat_qq() +
  stat_qq_line()+
  theme_bw()+
  scale_color_manual(values = my_colours) +
  facet_wrap(.~model, ncol=3)

```

### **Time series - annual max**

```{r message=FALSE}

historical_max_y <- historical_means %>%
  group_by(Yf, model) %>%
  dplyr::summarise(max=max(mean, na.rm=T))
```


```{r}

ggplot(historical_max_y) +
      geom_line(aes(x = as.numeric(Yf), y=max,
              color=model)) +
    theme_bw() + xlab("Year (Historical 1980 - 2000)") +
    ylab("Annual max of mean daily max temp (tasmax) oC") +
  scale_fill_brewer(palette = "Paired", name = "") +
   scale_colour_brewer(palette = "Paired", name = "") +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```


### **boxplot - annual max**


```{r}

historical_max_y %>%
  mutate(model = fct_reorder(model, max, .fun='median')) %>%
    ggplot(aes(x=reorder(model, max), y=max, fill=model)) +
    geom_boxplot() + theme_bw() +
    ylab("Annual max of mean daily max temp (tasmax) oC") + xlab("model") +
   scale_fill_manual(values = my_colours) +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```

The daily max is quite different than the means - something to bear in mind but interesting to think about - eg Run 4 here has the 2nd lowest spread of max max temp, but is selected above based on means


### **ANOVA**

Daily means:

```{r}
#-1 removes the intercept to compare coefficients of all Runs
av1 <- aov(mean ~ model - 1, historical_means)
av1$coefficients[order(av1$coefficients)]
```

Annual means:

```{r}
av2 <- aov(mean.annual ~ model - 1, historical_means_y)
av2$coefficients[order(av2$coefficients)]
```

Max of means:

```{r}

av3 <- aov(max ~ model - 1, historical_max_y)
av3$coefficients[order(av3$coefficients)]

```

Max vals are different but based on means then selection would be Run 02 (2nd lowest), Run 04 & Run 03, and Run 11 (2nd lowest)


### **2b. Y2020 - Y2040**

```{r}

Y <- rep(c(2021:2040), each=360)


dfs_Y21_Y40 <- lapply(names_Y21_Y40, function(i){
  df <- dfs_Y21_Y40[[i]]
  names(df) <- c("day", "mean", "sd")
  df$model <- i
  df$dn <- 1:nrow(df)
  df$Y <- Y
  return(df)
})

#Create a single df in long form of Runs for the Y21_Y40 period
Y21_Y40_means <- dfs_Y21_Y40 %>% reduce(rbind)

```

### **Time series - daily**

```{r}

ggplot(Y21_Y40_means) +
    geom_line(aes(x=dn, y=mean, group=model, colour=model)) +
  # Removing sd ribbon for ease of viewing
  #geom_ribbon(aes(x =dn, ymin = mean - sd, ymax= mean + sd), alpha=0.4) +
    theme_bw() + xlab("Daily (1980 - 2000)") +
    ylab("Daily mean max temp (tasmax) oC") +
  #scale_fill_brewer(palette = "Paired", name = "") +
   scale_colour_brewer(palette = "Paired", name = "") +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "none") +
  facet_wrap(.~ model, ncol=3) + guides(fill = FALSE)


```


### **boxplot - mean Y21_Y40**

```{r}
#Create a pallete specific to the runs so when reordered maintain the same colours
Y21_Y40_means$model <- as.factor(Y21_Y40_means$model)
c <- brewer.pal(12, "Paired")
my_colours <- setNames(c, levels(Y21_Y40_means$model))

```


```{r}

Y21_Y40_means %>%
  mutate(model = fct_reorder(model, mean, .fun='median')) %>%
    ggplot(aes(x=reorder(model, mean), y=mean, fill=model)) +
    geom_boxplot() + theme_bw() +
    ylab("Mean daily max temp (tasmax) oC") + xlab("model") +
    scale_fill_manual(values = my_colours) +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```

### **qqplot - daily means**

```{r fig.height=8}

ggplot(Y21_Y40_means, aes(sample=mean, colour=factor(model))) +
  stat_qq() +
  stat_qq_line()+
  theme_bw()+
  scale_color_manual(values = my_colours) +
  facet_wrap(.~model, ncol=3)

```


### **Time series - annual mean**

```{r message= FALSE, warning=FALSE}

#Aggregating to year for annual average

Y21_Y40_means$Yf <- as.factor(Y21_Y40_means$Y)

Y21_Y40_means_y <- Y21_Y40_means %>%
  group_by(Yf, model) %>%
  dplyr::summarise(mean.annual=mean(mean, na.rm=T), sd.annual=sd(mean, na.rm = T))
```

```{r}
ggplot(Y21_Y40_means_y) +
    geom_line(aes(x = as.numeric(Yf), y=mean.annual,
              color=model)) +
    theme_bw() + xlab("Year (2021 - 2040)") +
    ylab("Annual mean of mean daily max temp (tasmax) oC") +
  scale_fill_brewer(palette = "Paired", name = "") +
   scale_colour_brewer(palette = "Paired", name = "") +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```


```{r}

# Plotting with SDs in geom_ribbon to see if anything wildely different
ggplot(Y21_Y40_means_y) +
    geom_ribbon(aes(as.numeric(Yf), y=mean.annual,
                               ymin = mean.annual - sd.annual,
                              ymax= mean.annual + sd.annual,
                    fill=model),  alpha=0.4) +
    geom_line(aes(x = as.numeric(Yf), y=mean.annual,
              color=model)) +
    theme_bw() + xlab("Year (2021 - 2040)") +
    ylab("Annual mean of mean daily max temp (tasmax) oC") +
  scale_fill_brewer(palette = "Paired", name = "") +
   scale_colour_brewer(palette = "Paired", name = "") +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + facet_wrap(.~model, ncol=3)

```

### **boxplot - annual mean 2021 - 2040**

```{r}

Y21_Y40_means_y %>%
  mutate(model = fct_reorder(model, mean.annual, .fun='median')) %>%
    ggplot(aes(x=reorder(model, mean.annual), y=mean.annual, fill=model)) +
    geom_boxplot() + theme_bw() +
    ylab("Annual max daily max temp oC") + xlab("model") +
   scale_fill_manual(values = my_colours) +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```


### **qqplot - annual means**

```{r fig.height=8}

ggplot(Y21_Y40_means_y, aes(sample=mean.annual, colour=factor(model))) +
  stat_qq() +
  stat_qq_line()+
  theme_bw()+
  scale_color_manual(values = my_colours) +
  facet_wrap(.~model, ncol=3)

```

### **Time series - annual max**

```{r message=FALSE}

Y21_Y40_max_y <- Y21_Y40_means %>%
  group_by(Yf, model) %>%
  dplyr::summarise(max=max(mean, na.rm=T))
```


```{r}

ggplot(Y21_Y40_max_y) +
      geom_line(aes(x = as.numeric(Yf), y=max,
              color=model)) +
    theme_bw() + xlab("Year (2021 - 2040)") +
    ylab("Annual max of mean daily max temp (tasmax) oC") +
  scale_fill_brewer(palette = "Paired", name = "") +
   scale_colour_brewer(palette = "Paired", name = "") +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```


### **boxplot - annual max**


```{r}

Y21_Y40_max_y %>%
  mutate(model = fct_reorder(model, max, .fun='median')) %>%
    ggplot(aes(x=reorder(model, max), y=max, fill=model)) +
    geom_boxplot() + theme_bw() +
    ylab("Annual max of mean daily max temp (tasmax) oC") + xlab("model") +
   scale_fill_manual(values = my_colours) +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```


### **ANOVA**

Daily means:

```{r}
#-1 removes the intercept to compare coefficients of all Runs
av1 <- aov(mean ~ model - 1, Y21_Y40_means)
av1$coefficients[order(av1$coefficients)]

```

Annual means:

```{r}
av2 <- aov(mean.annual ~ model - 1, Y21_Y40_means_y)
av2$coefficients[order(av2$coefficients)]

```

Max of means

```{r}

av3 <- aov(max ~ model - 1, Y21_Y40_max_y)
av3$coefficients[order(av3$coefficients)]

```

Based on means then selection would be Run 02 (2nd lowest), Run 04 & Run 03, and Run 11 (2nd lowest)

Based on this period, the seelction would be: Run 05, Run 03, Run 08, Run 06 (so definetly Run 3 but others to be discussed)




### **2c. Y2061 - Y2080**

```{r}

Y <- rep(c(2061:2080), each=360)


dfs_Y61_Y80 <- lapply(names_Y61_Y80, function(i){
  df <- dfs_Y61_Y80[[i]]
  names(df) <- c("day", "mean", "sd")
  df$model <- i
  df$dn <- 1:nrow(df)
  df$Y <- Y
  return(df)
})

#Create a single df in long form of Runs for the Y61_Y80 period
Y61_Y80_means <- dfs_Y61_Y80 %>% reduce(rbind)

```

### **Time series - daily**

```{r}

ggplot(Y61_Y80_means) +
    geom_line(aes(x=dn, y=mean, group=model, colour=model)) +
  # Removing sd ribbon for ease of viewing
  #geom_ribbon(aes(x =dn, ymin = mean - sd, ymax= mean + sd), alpha=0.4) +
    theme_bw() + xlab("Day (2060 - 2080)") +
    ylab("Daily mean max temp (tasmax) oC") +
  #scale_fill_brewer(palette = "Paired", name = "") +
   scale_colour_brewer(palette = "Paired", name = "") +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "none") +
  facet_wrap(.~ model, ncol=3) + guides(fill = FALSE)


```


### **boxplot - mean Y61_Y80**

```{r}
#Create a pallete specific to the runs so when reordered maintain the same colours
Y61_Y80_means$model <- as.factor(Y61_Y80_means$model)
c <- brewer.pal(12, "Paired")
my_colours <- setNames(c, levels(Y61_Y80_means$model))

```


```{r}

Y61_Y80_means %>%
  mutate(model = fct_reorder(model, mean, .fun='median')) %>%
    ggplot(aes(x=reorder(model, mean), y=mean, fill=model)) +
    geom_boxplot() + theme_bw() +
    ylab("Mean daily max temp (tasmax) oC") + xlab("model") +
    scale_fill_manual(values = my_colours) +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```

### **qqplot - daily means**

```{r fig.height=8}

ggplot(Y61_Y80_means, aes(sample=mean, colour=factor(model))) +
  stat_qq() +
  stat_qq_line()+
  theme_bw()+
  scale_color_manual(values = my_colours) +
  facet_wrap(.~model, ncol=3)

```


### **Time series - annual mean**

```{r message= FALSE, warning=FALSE}

#Aggregating to year for annual average

Y61_Y80_means$Yf <- as.factor(Y61_Y80_means$Y)

Y61_Y80_means_y <- Y61_Y80_means %>%
  group_by(Yf, model) %>%
  dplyr::summarise(mean.annual=mean(mean, na.rm=T), sd.annual=sd(mean, na.rm = T))
```

```{r}
ggplot(Y61_Y80_means_y) +
    geom_line(aes(x = as.numeric(Yf), y=mean.annual,
              color=model)) +
    theme_bw() + xlab("Year (2061 - 2080)") +
    ylab("Annual mean of mean daily max temp (tasmax) oC") +
  scale_fill_brewer(palette = "Paired", name = "") +
   scale_colour_brewer(palette = "Paired", name = "") +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```


```{r}

# Plotting with SDs in geom_ribbon to see if anything wildely different
ggplot(Y61_Y80_means_y) +
    geom_ribbon(aes(as.numeric(Yf), y=mean.annual,
                               ymin = mean.annual - sd.annual,
                              ymax= mean.annual + sd.annual,
                    fill=model),  alpha=0.4) +
    geom_line(aes(x = as.numeric(Yf), y=mean.annual,
              color=model)) +
    theme_bw() + xlab("Year (2061 - 2080)") +
    ylab("Annual mean of mean daily max temp (tasmax) oC") +
  scale_fill_brewer(palette = "Paired", name = "") +
   scale_colour_brewer(palette = "Paired", name = "") +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + facet_wrap(.~model, ncol=3)

```

### **boxplot - annual mean Y61_Y80**

```{r}

Y61_Y80_means_y %>%
  mutate(model = fct_reorder(model, mean.annual, .fun='median')) %>%
    ggplot(aes(x=reorder(model, mean.annual), y=mean.annual, fill=model)) +
    geom_boxplot() + theme_bw() +
    ylab("Annual max daily max temp oC") + xlab("model") +
   scale_fill_manual(values = my_colours) +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```


### **qqplot - annual means**

```{r fig.height=8}

ggplot(Y61_Y80_means_y, aes(sample=mean.annual, colour=factor(model))) +
  stat_qq() +
  stat_qq_line()+
  theme_bw()+
  scale_color_manual(values = my_colours) +
  facet_wrap(.~model, ncol=3)

```

### **Time series - annual max**

```{r message=FALSE}

Y61_Y80_max_y <- Y61_Y80_means %>%
  group_by(Yf, model) %>%
  dplyr::summarise(max=max(mean, na.rm=T))
```


```{r}

ggplot(Y61_Y80_max_y) +
      geom_line(aes(x = as.numeric(Yf), y=max,
              color=model)) +
    theme_bw() + xlab("Year (2061 - 2080)") +
    ylab("Annual max of mean daily max temp (tasmax) oC") +
  scale_fill_brewer(palette = "Paired", name = "") +
   scale_colour_brewer(palette = "Paired", name = "") +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```


### **boxplot - annual max**


```{r}

Y61_Y80_max_y %>%
  mutate(model = fct_reorder(model, max, .fun='median')) %>%
    ggplot(aes(x=reorder(model, max), y=max, fill=model)) +
    geom_boxplot() + theme_bw() +
    ylab("Annual max of mean daily max temp (tasmax) oC") + xlab("model") +
   scale_fill_manual(values = my_colours) +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```


### **ANOVA**

Daily means:

```{r}
#-1 removes the intercept to compare coefficients of all Runs
av1 <- aov(mean ~ model - 1, Y61_Y80_means)
av1$coefficients[order(av1$coefficients)]

```

Annual means:

```{r}
av2 <- aov(mean.annual ~ model - 1, Y61_Y80_means_y)
av2$coefficients[order(av2$coefficients)]

```

Max of means

```{r}

av3 <- aov(max ~ model - 1, Y61_Y80_max_y)
av3$coefficients[order(av3$coefficients)]

```

Runs suggested by this slice are Run 05, Run 09, Run 03 and Run 11

Run 3 and 5 suggested above

## **3. Everything combined**

The result per time slice suggest different runs, aside from run 5

### Add in infill data

**Update 13.05.23** - Adding in the infill data, and taking the anova result across the whole time period

```{r infill means}

Y <- rep(c(2001:2020), each=360)

dfs_Y00_Y20 <- lapply(names_Y00_Y20, function(i){
  df <- dfs_Y00_Y20[[i]]
  names(df) <- c("day", "mean", "sd")
  df$model <- i
  df$dn <- 1:nrow(df)
  df$Y <- Y
  df$Yf <- as.factor(df$Y)
  return(df)
})


Y <- rep(c(2041:2060), each=360)

dfs_Y41_Y60 <- lapply(names_Y41_Y60, function(i){
  df <- dfs_Y41_Y60[[i]]
  names(df) <- c("day", "mean", "sd")
  df$model <- i
  df$dn <- 1:nrow(df)
  df$Y <- Y
  df$Yf <- as.factor(df$Y)
  return(df)
})


#Create a single df in long form as above
Y00_Y20_means <- dfs_Y00_Y20 %>% reduce(rbind)
Y41_Y60_means <- dfs_Y41_Y60 %>% reduce(rbind)
```

Assessing what the combined times slices suggest via anova

#### Daily means:

```{r}
#-1 removes the intercept to compare coefficients of all Runs
all.means <- rbind(historical_means, Y00_Y20_means, Y21_Y40_means, Y41_Y60_means, Y61_Y80_means)

x <- as.character(all.means$model)
all.means$model <- substr(x, nchar(x)-4, nchar(x))


av1 <- aov(mean ~ model - 1, all.means)
av1$coefficients[order(av1$coefficients)]
```

#### Annual means:

```{r}
# As above, creating annual means
infill.L <- list(Y00_Y20_means, Y41_Y60_means)

infill.L_y <- lapply(infill.L, function(x){
  means_y <- x %>%
  group_by(Yf, model) %>%
  dplyr::summarise(mean.annual=mean(mean, na.rm=T), sd.annual=sd(mean, na.rm = T))})

```

```{r}
all.means_y <- rbind(historical_means_y,
                     infill.L_y[[1]],
                     Y21_Y40_means_y,
                     infill.L_y[[2]],
                     Y61_Y80_means_y)

x <- as.character(all.means_y$model)
all.means_y$model <- substr(x, nchar(x)-4, nchar(x))

av2 <- aov(mean.annual ~ model - 1, all.means_y)
av2$coefficients[order(av2$coefficients)]

```


**Updated June 13th 2023 result**

Considering all together, suggests: Runs 05, Run07, Run08 and Run06
